Data_Setup

### 2020-06-12 R/Draft_figures_2020-06-02.Rmd
Make phenotype data for GWAS and add to Data_setup.Rmd

#

```{r}
metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_),
         LATITUDE = abs(LATITUDE),
         LONGITUDE = abs(LONGITUDE)) 
phe_all <- read_rds("../data/all_phenotype_meas_4wcr_gwas.rds")
sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")
phe_all$SUBPOP <- factor(phe_all$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
sites <- sites %>%
  mutate(manu_site = case_when(State != "TX" ~ State,
                               SITE == "KING" ~ "TX1",
                               SITE == "PKLE" ~ "TX2",
                               SITE == "TMPL" ~ "TX3",
                               SITE == "OVTN" ~ "TX4"))
sites$manu_site <- factor(sites$manu_site, levels = c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD"))

subpops <- enframe(unique(metadata$SUBPOP), name = NULL, value = "SUBPOP") %>%
  mutate(color_code = case_when(SUBPOP == "Atlantic" ~ "#6E91CB",
                                SUBPOP == "Gulf" ~ "#F47F72",    
                                SUBPOP == "Midwest" ~ "#442C83",
                                SUBPOP == "4X" ~ "grey", 
                                SUBPOP == "8X" ~ "#090906"))
subpops$SUBPOP <- factor(subpops$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))

>>>>

#
```{r}
metadata <- read_rds("../data/metadata.rds")
subpops <- read_rds("../data/subpop_color_coding.rds")
phenotypes <- 
  read_rds("../data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
sites <- read_rds("../data/sites.rds")
```
```{r}
phe_df <- phenotypes %>%
  filter(WHERE == "GWAS" & !(SITE %in% c("OVTN", "FRMI"))) %>%
  unite(col = "SITE_PHE", SITE, PHE) %>%
  filter(SUBPOP %in% c("Atlantic", "Midwest", "Gulf")) %>%
  select(-PLOT_GL, -WHERE, -YEAR) %>%
  pivot_wider(names_from = SITE_PHE, values_from = MEAS, values_fn = mean)

fwcr_df <- phenotypes %>%
  filter(WHERE == "FWCR" & !(SITE %in% c("OVTN", "FRMI", "MNHT")) & YEAR == 2019) %>%
  unite(col = "SITE_PHE", SITE, PHE) %>%
  select( -WHERE, -SUBPOP, -PLOT_GL) %>%
  pivot_wider(names_from = SITE_PHE, values_from = MEAS, values_fn = mean)

phe_df %>%
  ggplot(aes(x = PKLE_CRAIN, TMPL_CRAIN)) + geom_point(aes(color = SUBPOP)) + geom_abline(slope = 1, linetype = 2) +
  scale_color_manual(values = subpops$color_code)

fwcr_df %>%
  ggplot(aes(x = PKLE_CRAIN, TMPL_CRAIN)) + geom_point() + geom_abline(slope = 1, linetype = 2)

fwcr_df %>%
  filter(PKLE_CGDD_12C > 1200)
fwcr_df %>%
  filter(KBSM_CGDD_12C < 200)

gddcor <- phe_df %>%
  select(matches("CRAIN$")) %>%
  cor(., use = "pairwise")

gddcor %>%
  as_tibble(rownames = "SITE") %>%
  pivot_longer(cols = -SITE, names_to = "SITE2", values_to = "COR") %>%
  ggplot(aes(x = SITE, y = SITE2)) + 
  geom_tile(aes(fill = COR)) +
  scale_fill_viridis(option = "B") +
  theme(legend.position = "right", axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

phe_df <- phe_df %>% select(PLANT_ID, SUBPOP, matches("GDD"), matches("GR50"), matches("FL50"), matches("CRAIN"), matches("PTT"), everything())

atl_df <- phe_df %>% filter(SUBPOP == "Atlantic") 
atl_df <- atl_df %>%
  select(names(which(colSums(!is.na(atl_df)) > 100)))
midw_df <- phe_df %>% filter(SUBPOP == "Midwest") 
midw_df <- midw_df %>%
  select(names(which(colSums(!is.na(midw_df)) > 100)))
gulf_df <- phe_df %>% filter(SUBPOP == "Gulf") 
gulf_df <- gulf_df %>%
  select(names(which(colSums(!is.na(gulf_df)) > 100)))
midwgulf_df <- phe_df %>% filter(SUBPOP %in% c("Gulf", "Midwest")) 
midwgulf_df <- midwgulf_df %>%
  select(names(which(colSums(!is.na(midwgulf_df)) > 100)))

```
I've messed up making CGDD for the fourway cross somehow. especially for KBSM. Ugh.

KBSM & LINC most similar. CLMB, BRKG, LINC, and KBSM (and less so STIL) very similar. All corr > 0.6 though.

```{r}
phe_all <- read_rds("../data/older/all_phenotype_meas_4wcr_gwas.rds")
phe_all$SUBPOP <- factor(phe_all$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))

### 2020-06-12 data-raw/Data_Setup.Rmd
Make phenotype data for GWAS and add to Data_setup.Rmd

# metadata (germplasm)
```{r}
metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_))
write_rds(metadata, "../data/metadata.rds")
fourway <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/0-data/Phenology_2019_Multiple_Environment.csv")
```

@@

```

>>>>

# metadata (germplasm)
```{r}
metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_),
         LATITUDE = abs(LATITUDE),
         LONGITUDE = abs(LONGITUDE))
write_rds(metadata, "../data/metadata.rds")
```

@@

```
# All phenotypes
```{r}
phe_all <- read_rds("../data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
phenology <- read_rds("../data/Phenology_phenotypes_4wcr_and_gwas.rds")
```

```{r}
phenology %>%
  pivot_wider(names_from = PHE, values_from = MEAS) %>%
  left_join(phe_all) %>%
  dplyr::select(-CGDD_12C_5d, -PTT_12C_5d) %>%
  pivot_longer(cols = EM1:CRAIN_5d, names_to = "PHE", values_to = "MEAS") %>%
  filter(!is.na(MEAS)) %>%
  group_by(PHE, WHERE, YEAR) %>%
  tally() %>% arrange(desc(YEAR))

phenology %>%
  pivot_wider(names_from = PHE, values_from = MEAS) %>%
  left_join(phe_all) %>%
  dplyr::select(-CGDD_12C_5d, -PTT_12C_5d) %>%
  pivot_longer(cols = EM1:CRAIN_5d, names_to = "PHE", values_to = "MEAS") %>%
  filter(!is.na(MEAS)) %>%
  write_rds("../data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
```

# Phenotypes for GWAS
```{r}
phenotypes <- 
  read_rds("../data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")

phe_df <- phenotypes %>%
  filter(WHERE == "GWAS" & !(SITE %in% c("OVTN", "FRMI"))) %>%
  unite(col = "SITE_PHE", SITE, PHE) %>%
  filter(SUBPOP %in% c("Atlantic", "Midwest", "Gulf")) %>%
  select(-PLOT_GL, -WHERE, -YEAR) %>%
  pivot_wider(names_from = SITE_PHE, values_from = MEAS, values_fn = mean) %>%
  select(PLANT_ID, SUBPOP, matches("GDD"), matches("GR50"), matches("FL50"),
         matches("CRAIN"), matches("PTT"), everything())

midwgulf_df <- phe_df %>% filter(SUBPOP %in% c("Gulf", "Midwest")) 
midwgulf_df <- midwgulf_df %>%
  select(names(which(colSums(!is.na(midwgulf_df)) > 100))) %>%
  select(-SUBPOP)
midw_df <- phe_df %>% filter(SUBPOP == "Midwest") 
midw_df <- midw_df %>%
  select(names(which(colSums(!is.na(midw_df)) > 100))) %>%
  select(-SUBPOP)
gulf_df <- phe_df %>% filter(SUBPOP == "Gulf") 
gulf_df <- gulf_df %>%
  select(names(which(colSums(!is.na(gulf_df)) > 100))) %>%
  select(-SUBPOP)
atl_df <- phe_df %>% filter(SUBPOP == "Atlantic") 
atl_df <- atl_df %>%
  select(names(which(colSums(!is.na(atl_df)) > 100))) %>%
  select(-SUBPOP)
midwatl_df <- phe_df %>% filter(SUBPOP %in% c("Atlantic", "Midwest")) 
midwatl_df <- midwatl_df %>%
  select(names(which(colSums(!is.na(midwatl_df)) > 100))) %>%
  select(-SUBPOP)
gulfatl_df <- phe_df %>% filter(SUBPOP %in% c("Gulf", "Atlantic")) 
gulfatl_df <- gulfatl_df %>%
  select(names(which(colSums(!is.na(gulfatl_df)) > 100))) %>%
  select(-SUBPOP)

phe_subsets <- list(Midwest_and_Gulf = midwgulf_df, Midwest = midw_df, Gulf = gulf_df, Atlantic = atl_df, Midwest_and_Atlantic = midwatl_df, Gulf_and_Atlantic = gulfatl_df)
write_rds(phe_subsets, "../data/Phenotype_Subsets_for_Subpop_GWAS.rds")
names(phe_subsets)
```
